---
title: "Examples"
execute: 
  warning: false
---

```{r}
#| echo: false
params <- list()
params$choose_continent <- "`r params$choose_continent`"
```

## Data

Load the `gapminder` data:

```{r}
#| warning: false
library(gapminder)
```

## Examples

This section includes code for the examples shown. These may differ slightly from the examples shown in the live demonstration.

### Example 1: Parameterized plots with `ggplot2`

<details>
<summary>See example</summary>

Let's make a plot of how the life expectancy has changed in Asia over time. We start by filtering the data:

```{r}
#| message: false
library(dplyr)
gapminder_asia <- gapminder |>
  filter(continent == "Asia")
```

We then make a plot:

```{r}
#| message: false
library(ggplot2)
ggplot(data = gapminder_asia) +
  geom_boxplot(
    mapping = aes(x = factor(year), y = lifeExp)
  ) +
  labs(
    x = NULL, y = "Life Expectancy (years)",
    title = "Life Expectancy in Asia"
  )
```

If we wanted to create the same plot but for Europe, what would we need to change?

* The value that the data is filtered on
* The text in the title

Let's create a variable for the continent:

```{r}
choose_continent <- "Asia"
```

and then use that value in our code: 

```{r}
# Filter data
gapminder_continent <- gapminder |>
  filter(continent == choose_continent)

# Plot
library(glue)
ggplot(data = gapminder_continent) +
  geom_boxplot(
    mapping = aes(x = factor(year), y = lifeExp)
  ) +
  labs(
    x = NULL, y = "Life Expectancy (years)",
    title = glue("Life Expectancy in {choose_continent}")
  )
```

This makes it easier to turn it into a function:

```{r}
continent_plot <- function(choose_continent, data = gapminder) {
  # Filter data
  gapminder_continent <- data |>
    filter(continent == choose_continent)
  # Plot
  g <- ggplot(data = gapminder_continent) +
    geom_boxplot(
      mapping = aes(x = factor(year), y = lifeExp)
    ) +
    labs(
      x = NULL, y = "Life Expectancy (years)",
      title = glue("Life Expectancy in {choose_continent}")
    )
  return(g)
}
```

Check it works:

```{r}
continent_plot("Europe")
```

</details>

### Example 2: Parameterized reports with Quarto

<details>
<summary>See example</summary>

Here's a file, `example.qmd`, we'll use as a template for our report:

::: {.callout-note title="example.qmd" icon=false}

````
---
title: "Gapminder Report on Life Expectancy in Asia"
execute: 
  echo: false
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a box plot of how the life expectancy in Asia has changed over time.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in Asia"
# Edit the code below with your plot function!
plot(1:10, 1:10)
```

````

:::

Let's start by editing the `example.qmd` file to add a document parameter for the continent.

````
---
title: "Gapminder Report on Life Expectancy in Asia"
execute: 
  echo: false
params:
  choose_continent: "Europe"
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a box plot of how the life expectancy in Asia has changed over time.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in Asia"
# Edit the code below with your plot function!
plot(1:10, 1:10)
```

````

Then, we add our plotting function, and pass in the parameter:

````
---
title: "Gapminder Report on Life Expectancy in Asia"
execute: 
  echo: false
params:
  choose_continent: "Europe"
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a box plot of how the life expectancy in Asia has changed over time.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in Asia"
continent_plot <- function(choose_continent, data = gapminder) {
  # Filter data
  gapminder_continent <- data |>
    filter(continent == choose_continent)
  # Plot
  g <- ggplot(data = gapminder_continent) +
    geom_boxplot(
      mapping = aes(x = factor(year), y = lifeExp)
    ) +
    labs(
      x = NULL, y = "Life Expectancy (years)",
      title = glue("Life Expectancy in {params$choose_continent}")
    )
  return(g)
}
continent_plot(params$choose_continent)
```

````

We can use the parameter multiple times e.g. to add a table

````
---
title: "Gapminder Report on Life Expectancy in Asia"
execute: 
  echo: false
params:
  choose_continent: "Europe"
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a box plot of how the life expectancy in Asia has changed over time.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in Asia"
continent_plot <- function(choose_continent, data = gapminder) {
  # Filter data
  gapminder_continent <- data |>
    filter(continent == choose_continent)
  # Plot
  g <- ggplot(data = gapminder_continent) +
    geom_boxplot(
      mapping = aes(x = factor(year), y = lifeExp)
    ) +
    labs(
      x = NULL, y = "Life Expectancy (years)",
      title = glue("Life Expectancy in {params$choose_continent}")
    )
  return(g)
}
continent_plot(params$choose_continent)
```

Adding a table of data:


```{{r}}
#| label: tbl-stats
#| tbl-cap: "Statistics in Asia"
library(tinytable)
gapminder |> 
  filter(continent == params$choose_continent) |> 
  group_by(year) |> 
  summarise(across(c(lifeExp, pop, gdpPercap), mean)) |> 
  tt()
```

````

</details>

### Example 3: Parameterized text with Quarto

<details>
<summary>See example</summary>

Add inline code:

The number of rows is `` `r "\u0060r nrow(gapminder)\u0060"` ``.

which displays as:

The number of rows is `r nrow(gapminder)`.

We can apply this to our document and use the parameters inline:

````
---
title: "Gapminder Report on Life Expectancy in Asia"
execute: 
  echo: false
params:
  choose_continent: "Europe"
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a box plot of how the life expectancy in `r params$choose_continent` has changed over time.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in Asia"
continent_plot <- function(choose_continent, data = gapminder) {
  # Filter data
  gapminder_continent <- data |>
    filter(continent == choose_continent)
  # Plot
  g <- ggplot(data = gapminder_continent) +
    geom_boxplot(
      mapping = aes(x = factor(year), y = lifeExp)
    ) +
    labs(
      x = NULL, y = "Life Expectancy (years)",
      title = glue("Life Expectancy in {params$choose_continent}")
    )
  return(g)
}
continent_plot(params$choose_continent)
```

````

Including in the title:

````
---
title: "Gapminder Report on Life Expectancy in `r params$choose_continent`"
execute: 
  echo: false
params:
  choose_continent: "Europe"
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a box plot of how the life expectancy in `r params$choose_continent` has changed over time.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in Asia"
continent_plot <- function(choose_continent, data = gapminder) {
  # Filter data
  gapminder_continent <- data |>
    filter(continent == choose_continent)
  # Plot
  g <- ggplot(data = gapminder_continent) +
    geom_boxplot(
      mapping = aes(x = factor(year), y = lifeExp)
    ) +
    labs(
      x = NULL, y = "Life Expectancy (years)",
      title = glue("Life Expectancy in {params$choose_continent}")
    )
  return(g)
}
continent_plot(params$choose_continent)
```

````

We edit the figure caption slightly differently:

````
---
title: "Gapminder Report on Life Expectancy in `r params$choose_continent`"
execute: 
  echo: false
params:
  choose_continent: "Europe"
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a box plot of how the life expectancy in `r params$choose_continent` has changed over time.

```{{r}}
#| label: fig-life-exp
#| fig-cap: !expr 'paste0("Life expectancy in ", params$choose_continent)'
continent_plot <- function(choose_continent, data = gapminder) {
  # Filter data
  gapminder_continent <- data |>
    filter(continent == choose_continent)
  # Plot
  g <- ggplot(data = gapminder_continent) +
    geom_boxplot(
      mapping = aes(x = factor(year), y = lifeExp)
    ) +
    labs(
      x = NULL, y = "Life Expectancy (years)",
      title = glue("Life Expectancy in {params$choose_continent}")
    )
  return(g)
}
continent_plot(params$choose_continent)
```

````

</details>

### Example 4: Rendering multiple reports

<details>
<summary>See example</summary>

Render one report using the `quarto` R package:

```{r}
#| eval: false
#| echo: true
library(quarto)
quarto_render(
  input = "example.qmd",
  execute_params = list(
    continent = "Europe"
  )
)
```

Use the `purrr` package to create multiple reports at the same time:

```{r}
#| eval: false
#| echo: true
library(purrr)
walk(
  .x = unique(gapminder$continent),
  .f = ~quarto_render(
    input = "example.qmd",
    execute_params = list(
      choose_continent = .x
    )
  )
)
```

Why is there only one report? We need to give each report a different name, to stop it being overwritten. Let's use `glue()` again, this time for the file names.

```{r}
#| eval: false
#| echo: true
walk(
  .x = unique(gapminder$continent),
  .f = ~quarto_render(
    input = "example.qmd",
    output_file = glue("report_{.x}.html"),
    execute_params = list(
      choose_continent = .x
    )
  )
)
```

To ensure the images aren't overwritten as well, you may need to add `embed-resources: true` to your Quarto document YAML:

```yaml
---
format:
  html:
    embed-resources: true
---
```

This keeps any plots created embedded within the HTML document, instead of in a separate folder.

</details>

### Example 5: Conditional content

<details>
<summary>See example</summary>

* Split out the plot code from the plot output into two code blocks.
* Remember to keep the figure caption with the plot output.
* The default format is `html`. Check what it looks like in `pdf` by specifying the format.
* The cross references don't work quite correctly out of the box. We have two code blocks with the same figure reference, even though one is hidden. Switch to div syntax.

````
---
title: "Gapminder Report on Life Expectancy in `r params$choose_continent`"
format: pdf
execute: 
  echo: false
params:
  choose_continent: "Europe"
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
library(plotly)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a box plot of how the life expectancy in `r params$choose_continent` has changed over time.

```{{r}}
continent_plot <- function(choose_continent, data = gapminder) {
  # Filter data
  gapminder_continent <- data |>
    filter(continent == choose_continent)
  # Plot
  g <- ggplot(data = gapminder_continent) +
    geom_boxplot(
      mapping = aes(x = factor(year), y = lifeExp)
    ) +
    labs(
      x = NULL, y = "Life Expectancy (years)",
      title = glue("Life Expectancy in {params$choose_continent}")
    )
  return(g)
}
plot_output <- continent_plot(params$choose_continent)
```

::: {#fig-life-exp fig-cap="`r paste0('Life expectancy in ', params$choose_continent)`"}

::: {.content-visible when-format="html"}

```{{r}}
ggplotly(plot_output)
```

:::

::: {.content-visible unless-format="html"}

```{{r}}
plot_output
```

:::

:::


````

Make further edits to add a sentence when the continent is Europe.

````
---
title: "Gapminder Report on Life Expectancy in `r params$choose_continent`"
execute: 
  echo: false
params:
  choose_continent: "Europe"
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
library(plotly)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a box plot of how the life expectancy in `r params$choose_continent` has changed over time.

```{{r}}
continent_plot <- function(choose_continent, data = gapminder) {
  # Filter data
  gapminder_continent <- data |>
    filter(continent == choose_continent)
  # Plot
  g <- ggplot(data = gapminder_continent) +
    geom_boxplot(
      mapping = aes(x = factor(year), y = lifeExp)
    ) +
    labs(
      x = NULL, y = "Life Expectancy (years)",
      title = glue("Life Expectancy in {params$choose_continent}")
    )
  return(g)
}
plot_output <- continent_plot(params$choose_continent)
```

::: {#fig-life-exp fig-cap="`r paste0('Life expectancy in ', params$choose_continent)`"}

::: {.content-visible when-format="html"}

```{{r}}
ggplotly(plot_output)
```

:::

::: {.content-visible unless-format="html"}

```{{r}}
plot_output
```

:::

:::

```{{r}}
latest_begin <- if (params$choose_continent == "Europe") "::: {.content-visible}"
latest_end <- if (params$choose_continent == "Europe") ":::"
```

`r "\u0060r latest_begin\u0060"`

This is the most recent data available.

`r "\u0060r latest_end\u0060"`


````

</details>

### Example 6: Parameterized report sections

<details>
<summary>See example</summary>

Create a .R file and put the function in it. You don't need to make any changes.

```{r}
#| eval: false
#| echo: true
#| filename: setup.R
continent_plot <- function(choose_continent, data = gapminder) {
  # Filter data
  gapminder_continent <- data |>
    filter(continent == choose_continent)
  # Plot
  g <- ggplot(data = gapminder_continent) +
    geom_boxplot(
      mapping = aes(x = factor(year), y = lifeExp)
    ) +
    labs(
      x = NULL, y = "Life Expectancy (years)",
      title = glue("Life Expectancy in {params$choose_continent}")
    )
  return(g)
}
```

The use `file:` to load it in.

````
---
title: "Gapminder Report on Life Expectancy in `r params$choose_continent`"
execute: 
  echo: false
params:
  choose_continent: "Europe"
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

```{{r}}
#| file: setup.R
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a box plot of how the life expectancy in `r params$choose_continent` has changed over time.

```{r}
#| label: fig-life-exp
#| fig-cap: !expr 'paste("Life expectancy in ", params$choose_continent)'
continent_plot(params$choose_continent)
```


````

Create a child document containing a section heading and the plot code.

::: {.callout-note title="_child.qmd" icon=false}

````

## `r params$choose_continent`

```{{r}}
continent_plot(params$choose_continent)
```

````

:::

Then call it in the main file with `knit_child()`, and check it works for one continent. Set output to `asis` so that Quarto renders the Markdown content correctly, and run it with `quiet = TRUE` to stop the rendering messages from coming through.

````
---
title: "Gapminder Report on Life Expectancy
execute: 
  echo: false
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

```{{r}}
#| file: setup.R
```

```{{r}}
#| output: asis
params <- list(choose_continent = "Europe")
a <- knitr::knit_child('_child.qmd', quiet = TRUE)
cat(a, sep='\n')
```

````

The we can loop over all continents.

````
---
title: "Gapminder Report on Life Expectancy
execute: 
  echo: false
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

```{{r}}
#| file: setup.R
```

```{{r}}
#| output: asis
all_continents <- as.character(unique(gapminder$continent))
for (i in seq_len(length(all_continents))) {
  params <- list(choose_continent = all_continents[i])
  a <- knitr::knit_child('_child.qmd', quiet = TRUE)
  cat(a, sep='\n')
}
```

````

</details>

