---
title: "Exercises"
execute: 
  eval: true
  echo: true
---

```{r}
#| echo: false
params <- list()
params$choose_year <- "`r params$choose_year`"
```

## Data

Load the `gapminder` data:

```{r}
#| warning: false
library(gapminder)
```

## Exercises

### Exercise 1: Parameterized plots with `ggplot2`

1. Create a variable to define a year in the `gapminder` data that you want to plot.

<details>
<summary>Solution</summary>

```{r}
choose_year <- 1952
```

</details>

2. Edit the code below to use this new year variable.

```{r}
#| eval: false
library(dplyr)
library(ggplot2)
# Median life expectancy per continent
gapminder_1952 <- gapminder |>
  filter(year == 1952) |>
  group_by(continent) |>
  summarise(
    median_lifeExp = median(lifeExp)
  )
# Bar chart
ggplot(data = gapminder_1952) +
  geom_col(
    mapping = aes(
      x = median_lifeExp,
      y = reorder(continent, median_lifeExp)
    )
  ) +
  labs(
    x = "Median Life Expectancy (years)", y = NULL,
    title = "Life Expectancy in 1952"
  )
```

<details>
<summary>Solution</summary>

```{r}
#| warning: false
library(dplyr)
library(ggplot2)
library(glue)

# Median life expectancy per continent
gapminder_year <- gapminder |>
  filter(year == choose_year) |>
  group_by(continent) |>
  summarise(
    median_lifeExp = median(lifeExp)
  )

# Bar chart
ggplot(data = gapminder_year) +
  geom_col(
    mapping = aes(
      x = median_lifeExp,
      y = reorder(continent, median_lifeExp)
    )
  ) +
  labs(
    x = "Median Life Expectancy (years)", y = NULL,
    title = glue("Life Expectancy in {choose_year}")
  )
```

</details>

3. Turn your code into a function that takes the data and year variable as arguments.

<details>
<summary>Solution</summary>

```{r}
barchart_year <- function(choose_year, data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      )
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
```

Check it works:

```{r}
barchart_year(2002)
```

</details>

4. **Bonus**: Add another argument to your function that allows a user to define the bar colour.

<details>
<summary>Solution</summary>

```{r}
barchart_year <- function(
    choose_year, fill = "blue", data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      ),
      fill = fill
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
```

Check it works:

```{r}
barchart_year(2002, fill = "purple")
```

</details>

5. **Bonus**: How could you improve your function?

<details>
<summary>Solution</summary>

* Not all year values are included in the data. Add a check that the value is a number that is included in the `gapminder` data. if it isn't, add a more useful message.
* Write some documentation!

</details>

### Exercise 2: Parameterized reports with Quarto

1. Edit the `report.qmd` file to add a parameter for year.

::: {.callout-warning title="report.qmd"}

````
---
title: "Gapminder Report: 1952"
execute: 
  echo: false
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year 1952 for each continent.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in 1952"
# Edit the code below with your plot function!
plot(1:10, 1:10)
```

````

:::

<details>
<summary>Solution</summary>

````
---
title: "Gapminder Report: 1952"
execute: 
  echo: false
params:
  choose_year: 1952
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year 1952 for each continent.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in 1952"
# Edit the code below with your plot function!
plot(1:10, 1:10)
```
````

</details>

2. Add your plotting function to the quarto document, and pass the parameter into the plotting function. Check it renders.

<details>
<summary>Solution</summary>

````
---
title: "Gapminder Report: 1952"
execute: 
  echo: false
params:
  choose_year: 1952
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year 1952 for each continent.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in 1952"
barchart_year <- function(
    choose_year, fill = "blue", data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      ),
      fill = fill
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
barchart_year(params$choose_year)
```

````
</details>

3. Change the value of the parameter, and re-render.


### Exercise 3: Parameterized text with Quarto

1. Edit the sentence `...a plot of the median life expectancy in the year 1952 for each continent...` to use inline code to add the document parameter for the year. 

<details>
<summary>Solution</summary>

````
---
title: "Gapminder Report: 1952"
execute: 
  echo: false
params:
  choose_year: 1952
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year `r params$choose_year` for each continent.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in 1952"
barchart_year <- function(
    choose_year, fill = "blue", data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      ),
      fill = fill
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
barchart_year(params$choose_year)
```

````

</details>

2. Use inline code for the title as well.

<details>
<summary>Solution</summary>

````
---
title: "Gapminder Report: `r params$choose_year`"
execute: 
  echo: false
params:
  choose_year: 1952
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year `r params$choose_year` for each continent.

```{{r}}
#| label: fig-life-exp
#| fig-cap: "Life expectancy in 1952"
barchart_year <- function(
    choose_year, fill = "blue", data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      ),
      fill = fill
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
barchart_year(params$choose_year)
```

````

</details>

3. Edit the figure caption to use inline code to add the document parameter for the year. 

<details>
<summary>Solution</summary>

````
---
title: "Gapminder Report: `r params$choose_year`"
execute: 
  echo: false
params:
  choose_year: 1952
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year `r params$choose_year` for each continent.

```{{r}}
#| label: fig-life-exp
#| fig-cap: !expr 'paste("Life expectancy in ", params$choose_year)'
barchart_year <- function(
    choose_year, fill = "blue", data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      ),
      fill = fill
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
barchart_year(params$choose_year)
```

````

</details>

### Exercise 4: Rendering multiple reports

1. Render a version of your report for the year `1972` using the `quarto_render()` function.

<details>
<summary>Solution</summary>

```{r}
#| eval: false
quarto_render(
  input = "report.qmd",
  execute_params = list(choose_year = 1972)
)
```

</details>

2. Render a version of your report for all years in the data.

<details>
<summary>Solution</summary>

```{r}
#| eval: false
walk(
  .x = unique(gapminder$year),
  .f = ~quarto_render(
    input = "report.qmd",
    output_file = glue("report_{.x}.html"),
    execute_params = list(
      choose_year = .x
    )
  )
)
```

</details>


### Exercise 5: Conditional content

1. Edit the .qmd file from Exercise 3. When the format is HTML show an interactive chart using `plotly`, otherwise show a static chart.

<details>
<summary>Solution</summary>

````
---
title: "Gapminder Report: `r params$choose_year`"
execute: 
  echo: false
params:
  choose_year: 1952
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
library(plotly)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year `r params$choose_year` for each continent.

```{{r}}
barchart_year <- function(
    choose_year, fill = "blue", data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      ),
      fill = fill
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
plot_output <- barchart_year(params$choose_year)
```

::: {#fig-life-exp fig-cap="`r paste0('Life expectancy in ', params$choose_year)`"}

::: {.content-visible when-format="html"}

```{{r}}
ggplotly(plot_output)
```

:::

::: {.content-visible unless-format="html"}

```{{r}}
plot_output
```

:::

:::


````

</details>

2. When the year parameter is 2007, add a sentence to state that it is the most recent data available.

<details>
<summary>Solution</summary>

````
---
title: "Gapminder Report: `r params$choose_year`"
execute: 
  echo: false
params:
  choose_year: 1952
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
library(plotly)
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year `r params$choose_year` for each continent.

```{{r}}
barchart_year <- function(
    choose_year, fill = "blue", data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      ),
      fill = fill
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
plot_output <- barchart_year(params$choose_year)
```

::: {#fig-life-exp fig-cap="`r paste0('Life expectancy in ', params$choose_year)`"}

::: {.content-visible when-format="html"}

```{{r}}
ggplotly(plot_output)
```

:::

::: {.content-visible unless-format="html"}

```{{r}}
plot_output
```

:::

:::

```{{r}}
latest_begin <- if (params$choose_year == 2007) "::: {.content-visible}"
latest_end <- if (params$choose_year == 2007) ":::"
```

`r "\u0060r latest_begin\u0060"`

This is the most recent data available.

`r "\u0060r latest_end\u0060"`


````

</details>

3. **Bonus**: Can you edit the logic you applied in question 2, to check programmatically if it's the last year in the data?

<details>
<summary>Solution</summary>

```{r}
#| eval: false
#| echo: true
last_year <- max(gapminder$year)
latest_begin <- if (params$choose_year == last_year) "::: {.content-visible}"
latest_end <- if (params$choose_year == last_year) ":::"
```


</details>

### Exercise 6: Parameterized report sections

1. Separate out the `barchart_year()` function into a separate R script.

2. Load it in using `file:` instead.

<details>
<summary>Solution</summary>

Create a .R file and put the function in it. You don't need to make any changes.

```{r}
#| eval: false
#| echo: true
#| filename: setup.R
barchart_year <- function(
    choose_year, fill = "blue", data = gapminder) {
  # Median life expectancy per continent
  gapminder_year <- gapminder |>
    filter(year == choose_year) |>
    group_by(continent) |>
    summarise(
      median_lifeExp = median(lifeExp)
    )
  # Bar chart
  g <- ggplot(data = gapminder_year) +
    geom_col(
      mapping = aes(
        x = median_lifeExp,
        y = reorder(continent, median_lifeExp)
      ),
      fill = fill
    ) +
    labs(
      x = "Median Life Expectancy (years)", y = NULL,
      title = glue("Life Expectancy in {choose_year}")
    )
  return(g)
}
```

The use `file:` to load it in.

````
---
title: "Gapminder Report: `r params$choose_year`"
execute: 
  echo: false
params:
  choose_year: 1952
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

```{{r}}
#| file: setup.R
```

This is a report using the [Gapminder](https://www.gapminder.org/data/) data. @fig-life-exp shows a plot of the median life expectancy in the year `r params$choose_year` for each continent.

```{r}
#| label: fig-life-exp
#| fig-cap: !expr 'paste("Life expectancy in ", params$choose_continent)'
barchart_year(params$choose_year)
```

````

</details>

3. Create a child file containing the year as a section heading, followed by the chart.

<details>
<summary>Solution</summary>

::: {.callout-note title="_child.qmd" icon=false}

````

## `r params$choose_year`

```{{r}}
barchart_year(params$choose_year)
```

````

:::

</details>

4. Render document with a section for each year in the data.

<details>
<summary>Solution</summary>

````
---
title: "Gapminder Report
execute: 
  echo: false
---

```{{r}}
#| label: load-pkgs
#| warning: false
library(gapminder)
library(dplyr)
library(ggplot2)
library(glue)
```

```{{r}}
#| file: setup.R
```

```{{r}}
#| output: asis
all_years <- as.character(unique(gapminder$year))
for (i in seq_len(length(all_years))) {
  params <- list(choose_year = all_years[i])
  a <- knitr::knit_child('_child.qmd', quiet = TRUE)
  cat(a, sep='\n')
}
```

````

</details>




